---
title: "Projections for the Covid-19 pandemic"
author: Raphael Volz (rv@volzinnovation.com)
output: bookdown::html_document2
params:
country: Germany
---

Please note that this project was created for one wave predictions only. **Since we have multiple waves of COVID-19 infections around the globe, our models currently provide no reliable predictions.** But, we are working on models for several waves of infections.

```{r , echo=F  , messages=F, warning=F, include=F}
library(stringr)
library(knitr)
library(DT)
library(dplyr)
p = read.csv("population.csv")
d = read.csv("models/history.csv")
```
        
**Disclaimer:** This is preliminary research provided on a timely basis for information only. It is not intended as commercial, legal, medical, financial or political advice.

# Interim results per country

```{r , messages=F, warning=F, echo=F}
last_d = d$date[nrow(d)]
d = subset(d, date == last_d)
d = merge(d, p, by.x="country", by.y="Country", sort = TRUE)
d = distinct(d, country,  .keep_all = TRUE)
filename = str_replace_all(d$country, c(" "="","\\*"="","US"="USA","'"=""))
d$link = paste0("[",d$country,"](https://www.volzinnovation.com/covid-19_SARS-CoV-2_corona/reports/latest/",filename,".html)")
kable(d$link, col.names=c("Country"))

```

# Frequently Asked Question (FAQ)

## What is this ?
An analysis of Covid-19 case time series around the world with projections and forecasts how cases numbers might grow in the week ahead due to the spread of SARS-CoV-2 (Corona) virus infections.

## How does this work ?

* We use the [Johns Hopkins University data set published on Github by the CSSE group](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/).
* The published reports are updated daily.
* [Models are open source](https://github.com/volzinnovation/covid-19_SARS-CoV-2_corona/tree/master/src) and can be run with the [free statistical software R](https://www.r-project.org/) on your system of choice.

## Whom should I contact for questions ?

Please contact Raphael Volz (rv@volzinnovation.com) in case of questions about the analyses. Your feedback is welcome!