```{r setup, include=FALSE}
# README FIRST
#
#
# Install required packages first, by uncommenting the following line...
# install.packages("forecast,knitr,minpack.lm,rmarkdown,zoo,xts,reshape,plyr,plotly, sicegar, lubridate, knitr, doBy")
# Import required libraries
#
#
#
#
# Use country from build script parameter, or default to a country (second parameter)
################# DEBUG WITH COUNTRY NAME *********************
DEBUG_COUNTRY = "US"
################# DEBUG WITH A COUNTRY NAME *********************
#
#
#
#
# Get country name from build script parameter
country = ifelse(exists("country"),country,DEBUG_COUNTRY) 

library(forecast)
library(knitr)
library(minpack.lm)
library(rmarkdown)
library(zoo)
library(xts)
library(reshape)
library(plyr)
library(plotly)
library(sicegar)
library(lubridate)
library(knitr)
library(doBy)
library(bookdown)
# Set formating of Time to German
# Sys.setlocale("LC_TIME", "de_DE.UTF-8")
# Set 3 decimal points
options(digits = 3)
# Set no scientific notation
options(scipen = 999)

# Change country name to adhere to the country rules
thecountry <- list("Bahamas","Cayman Islands", "Central African Republic", "Comoros", "Czech Republic", 
"Dominican Republic", "Falkland Islands", "Gambia", "Isles of Man", "Ivory Coast", 
"Leeward Islands", "Maldives", "Marshall Islands", "Netherlands", "Netherlands Antilles", "Philippines", "Solomon Islands", 
"Turks and Caicos Islands", "United Arab Emirates", "United Kingdom", "US","Virgin Islands")
# change countryname for use in the text
countryname = ifelse(country %in% thecountry, paste0("the ", country),country)
countryname = ifelse(country == "Korea, South", "South Korea",country)

# Automated assessment of reliability
assessment = TRUE

# Load Data from Johns Hopkins Github Repository
confirmed = read.csv('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv')


# Calculate last date reported and reference dates
columns = names(confirmed)
yesterday = columns[length(columns)]
yesterday = substr(yesterday,2,nchar(yesterday))
yesterday = as.Date(strptime(yesterday, "%m.%d.%y"))
today = yesterday+1
tomorrow = today+1
oneweek = today+7
lastweek = today-7

# Transform data to sensible timeseries format
confirmed_t = melt(confirmed, id=c("Province.State","Country.Region","Lat","Long"))
names(confirmed_t) = c("State","Country","Lat","Lon","Date","Confirmed") # Proper Col Names
confirmed_t$Date = as.character(confirmed_t$Date) # To Character
confirmed_t$Date = substr(confirmed_t$Date,2,nchar(confirmed_t$Date)) # Drop X
confirmed_t$Date = as.Date(strptime(confirmed_t$Date, "%m.%d.%y")) # To Date

# Select data relevant to country
DE_ALL = subset(confirmed_t, Country==country )
# Summarize by regions, where available, e.g. US
DE = summaryBy(Confirmed ~ Date, data=DE_ALL, FUN = sum)
# Look only to data when some cases are confirmed
DE = subset(DE, DE$Confirmed > 0)
# Last known cases
last_confirmed = DE$Confirmed[nrow(DE)]
# Create time series
z = zoo(DE$Confirmed, DE$Date)
# Deduplicate time series
z = aggregate(z, index,  tail, 1)
#
#
# Create Gompertz model
#
#
# treat time as a number instead of dates
time = seq(1,nrow(DE)) 
# take intensity from confirmed infections
intensity = DE$Confirmed
# reverse the time vector for weighting newer data (UNUSED)
t2 = rev(time)
# create a small data frame for training
train = data.frame(time, intensity, t2)
# Fit model to data iteratvely
gomp <- nlsLM(intensity ~ a * exp( -b * exp( - c* time ) ),
                data = train,
                #weights = wfct(1/t2), # prefer newer data
                control = list(maxiter = 1024),
                start = list(a = 41000000, b = 40,c = 0.1),
                upper = c(a = 1000000000, b = 10000, c = 10 ),
                lower = c(a = 0, b=1, c=0.0000001) )
# Make a prediction using the model formula and estimate parameters
new_gomp = gomp$m$getAllPars()["a"] * exp( - gomp$m$getAllPars()["b"] * exp( - gomp$m$getAllPars()["c"] * time ))

# create comparison week ago model using the data available last week
DE_weekago = subset(DE, Date <= lastweek)

# Create comparison model using data until last week, see comments above
time = seq(1, nrow(DE_weekago))
intensity = DE_weekago$Confirmed
t2 = rev(time)
train = data.frame(time, intensity, t2)
gomp_ago <- nlsLM(intensity ~ a * exp( -b * exp( - c* time ) ), 
                data = train,
                #weights = wfct(1/t2),
                control = list(maxiter = 1024),
                start = list(a = 30000000, b = 40,c = 0.1),
                upper = c(a = 82000000, b = 10000, c = 10 ),
                lower = c(a = 0, b=1, c=0.0000001) )
time = seq(1,nrow(DE))
p_gomp_ago = gomp_ago$m$getAllPars()["a"] * exp( - gomp_ago$m$getAllPars()["b"] * exp( - gomp_ago$m$getAllPars()["c"] * time ))

# Calculate daily diffs
z = aggregate(z, index,  tail, 1) # Deduplicate zoo
delta = diff(z)
# Determine last delta
last_delta = coredata(delta)[length(coredata(delta))]
# Calculate moving average with window of 5 days (as RKI does)
trend = ma(delta, order = 5, centre = F)
#
# Estimate series of R0 values (as RKI does)
R0 = lag(trend, 4) / trend
R0_df = data.frame(R0 = coredata(R0))
R0_df = subset(R0_df, !is.na(R0_df$R0))
last_R0 = R0_df$R0[nrow(R0_df)]
#
# Compute relative increase
Increase =  (diff(z)/lag(z, 1))*100
Increase_d = coredata(Increase)
# Compute last relative increase in percent
last_increase_percent = round(Increase_d[length(Increase_d)],digits=1)
# Compute days of doubling infections
double_days = log(2)/ log ( 1 + last_increase_percent ) *100

#
#
# Compute double sigmoidal model
#
# Compute
time <- seq(1, length(delta))
# Names are fixed in sicegar package
# Train on deltas
dataInput <- data.frame(intensity = coredata(delta), time = time)
# Normalize data to 0..1
normalizedInput <- sicegar::normalizeData(dataInput, dataInputName = "sample001")
# Find a fit for the double sigmoidal function
doubleSigmoidalModel <- sicegar::multipleFitFunction(dataInput = normalizedInput,
                                                     model = "doublesigmoidal",
                                                     n_runs_min = 20,
                                                     n_runs_max = 500,
                                                     showDetails = FALSE)
# Obtain model parameters
doubleSigmoidalModel <- sicegar::parameterCalculation(doubleSigmoidalModel)
# Use the standard rendering (TODO switch to custom plot)
#fig01 <- sicegar::figureModelCurves(dataInput = normalizedInput,
#                                    doubleSigmoidalFitVector = doubleSigmoidalModel,
#                                    showParameterRelatedLines = TRUE)
# Render figure in document
#print(fig01)
#
# Obtain predictions
#
# Relate to all available data
time <- seq(1, nrow(DE))
# Make prediction using model
new_dsm <- sicegar::doublesigmoidalFitFormula(time,
                  finalAsymptoteIntensityRatio = doubleSigmoidalModel$finalAsymptoteIntensityRatio_Estimate,
                  maximum = doubleSigmoidalModel$maximum_y,
                  slope1Param = doubleSigmoidalModel$slope1Param_Estimate,
                  midPoint1Param = doubleSigmoidalModel$midPoint1Param_Estimate,
                  slope2Param = doubleSigmoidalModel$slope2Param_Estimate,
                  midPointDistanceParam = doubleSigmoidalModel$midPointDistanceParam_Estimate)


# Cumulate model prediction to show aside cumulative figures
new_dsm_cum = cumsum(new_dsm)

# Create comparison model using data until last week, see comments above
# Create time series for week old data
z_ago = zoo(DE_weekago$Confirmed, DE_weekago$Date)
z_ago = aggregate(z_ago, index, tail, 1) # Deduplicate time series
delta_ago = diff(z_ago)
time_ago <- seq(1, length(delta_ago))
# Names are fixed in sicegar package
# Train on deltas
dataInput_ago <- data.frame(intensity = coredata(delta_ago), time = time_ago)
# Normalize data to 0..1
normalizedInput_ago <- sicegar::normalizeData(dataInput_ago, dataInputName = "weekago")
# Find a fit for the double sigmoidal function
dsm_ago <- sicegar::multipleFitFunction(dataInput = normalizedInput_ago,
                                                     model = "doublesigmoidal",
                                                     n_runs_min = 20,
                                                     n_runs_max = 500,
                                                     showDetails = FALSE)
# Obtain model parameters
dsm_ago <- sicegar::parameterCalculation(dsm_ago)
# Assign values to variables (should have done that much earlier, TODO update uses to those names)
t_d = round(doubleSigmoidalModel$midPoint2_x, digits=0)
t_i = round(doubleSigmoidalModel$midPoint1_x, digits=0)
a_d = round(doubleSigmoidalModel$slope2Param_Estimate, digits=2)
a_i = round(doubleSigmoidalModel$slope1Param_Estimate, digits=2)
t_max = round(doubleSigmoidalModel$maximum_x, digits=0)
I_final_d = round(doubleSigmoidalModel$endDeclinePoint_y, digits=0)
I_max_d = round(doubleSigmoidalModel$reachMaximum_y, digits=0)
I_max_c = gomp$m$getAllPars()["a"]
beta_c = gomp$m$getAllPars()["b"]
alpha_c = gomp$m$getAllPars()["c"]
# Check whether a maximum can be found for the DSM model
I_max_dsm = ifelse( ( a_d > 0 ) && ( I_final_d < 0.5 * I_max_d ),  
                    new_dsm_cum[ t_max ] +  2 * ( new_dsm_cum[t_d] - new_dsm_cum[ t_max ]) , -1 ) # not really fixed asymptote, -1 if there is no recovery yet
# Assess model
if(t_d <= t_max) { assessment = FALSE} # Decay is before maximum
if(t_d >= nrow(DE) ) { assessment = FALSE } # Decay starts in the future
if(a_d >= 1  ) { assessment = FALSE } # Decay is too high

# find first date of infections
min_d = min(DE$Date)
# create a sequence of dates to create models for
dates = seq((min_d + 7),( today - 1),"day")
# create empty models2 dataframe to observe convergence
models2 = data.frame()
# iterate over data
for (date in dates) {
 # Contine with next date when a fit cannot be found for some reason
 try({
  # Subset
  DE_weekago = subset(DE, Date <= date)
  # Create time span for the model
  time = seq(1, nrow(DE_weekago))
  # reverse time for weighting (UNUSED)
  t2 = rev(time)
  # Create intensity of from Confirmed data
  intensity = DE_weekago$Confirmed
  # Create training data set
  train = data.frame(time, intensity, t2)
  # Fit Gompertz model to training data
  gomp <- nlsLM(intensity ~ a * exp( -b * exp( - c* time ) ), 
                data = train,
                #weights = wfct(1/t2),
                control = list(maxiter = 1024),
                start = list(a = 41000000, b = 40,c = 0.1),
                #weights = wfct(1/intensity^2),
                upper = c(a = 1000000000, b = 10000, c = 10 ),
                lower = c(a = 0, b=1, c=0.0000001) )
  # Create test set
  DE_test = subset(DE, Date > date)
  # Create future time sequence for prediction
  time_t = seq(length(time)+1, length(time)+nrow(DE_test))
  # Reverse time for weighting (UNUSED)
  t2_t = rev(time_t)
   # Assign intensity from Confirmed infections
  intensity_t = DE_test$Confirmed
  # Make predictions using Gompertz model
  p_gomp = gomp$m$getAllPars()["a"] * exp( - gomp$m$getAllPars()["b"] * exp( - gomp$m$getAllPars()["c"] * time_t ))
  # Calculate Test Set RMSE
  RMSE_gomp = sqrt(sum((intensity_t- p_gomp)^2)/nrow(DE_test))
  # Create new record
  m2 = data.frame(Date=c(date), 
                  KnownCases=c(DE_weekago$Confirmed[nrow(DE_weekago)]),   
                  Rows=nrow(DE_weekago),
                  Maximum=c(gomp$m$getAllPars()["a"]),
                  B=c(gomp$m$getAllPars()["b"]),
                  C=c(gomp$m$getAllPars()["c"]),
                  RMSE_train=c(sqrt( sum ( (DE_weekago$Confirmed - gomp$m$fitted() )^2 ) / nrow(train))),
                 RMSE_test = c(RMSE_gomp),
                 pred_delta_abs = c( gomp$m$getAllPars()["a"] * exp( - gomp$m$getAllPars()["b"] * exp( - gomp$m$getAllPars()["c"] * nrow(DE) ))  ))
  # Add record to models
  models2  = rbind.fill(models2,m2)
})
}
# Convert Date column to Date type
models2$Date=as.Date(models2$Date, origin="1970-01-01")

# Create prediction data set from model(s)
comparison = data.frame(Date = models2$Date, 
                        actual=DE$Confirmed[nrow(DE)],
                        gomp=models2$pred_delta_abs)
# Check convergence
convergent = 0 # 0 means did not yet converge
stop = F
criteria = 2 # nowcast should be maximum 2% off the real value
# search backwards
for(i in rev( seq(1,nrow(models2) ) ) ) {
    if (
        (( comparison$gomp[i] >= (1 - criteria/100) * DE$Confirmed[nrow(DE)] ) 
        && ( comparison$gomp[i] <= (1 + criteria/100) * DE$Confirmed[nrow(DE)] ))
        ) {
        # row still in search criteria
        #cat(paste0(i," : ", (1 - criteria/100) * DE$Confirmed[nrow(DE)] , "<=", comparison$gomp[i], "<=" ,  (1 + criteria/100) * DE$Confirmed[nrow(DE)] , "\n"))
        convergent = i
    } else {
      #cat(paste0("STOP"," : ", (1 - criteria/100) * DE$Confirmed[nrow(DE)] , "<=", comparison$gomp[i], "<=" ,  (1 + criteria/100) * DE$Confirmed[nrow(DE)] , "\n"))
      # stop searching
      break  
    }
}

# Create new data for next week
Date = seq(today,oneweek, by="days")
time = seq(nrow(DE)+1,nrow(DE)+8)
# Gompertz Model
p_gomp = gomp$m$getAllPars()["a"] * exp( - gomp$m$getAllPars()["b"] * exp( - gomp$m$getAllPars()["c"] * time ))
# Double sigmoidal model
# Obtain predictions
p_dsm <- sicegar::doublesigmoidalFitFormula(time,
                  finalAsymptoteIntensityRatio = doubleSigmoidalModel$finalAsymptoteIntensityRatio_Estimate,
                  maximum = doubleSigmoidalModel$maximum_y,
                  slope1Param = doubleSigmoidalModel$slope1Param_Estimate,
                  midPoint1Param = doubleSigmoidalModel$midPoint1Param_Estimate,
                  slope2Param = doubleSigmoidalModel$slope2Param_Estimate,
                  midPointDistanceParam = doubleSigmoidalModel$midPointDistanceParam_Estimate)
dsm_cum = cumsum(p_dsm) + max(DE$Confirmed)

# Create prediction data frame with Date and Models
Prediction = data.frame(Date,
                        Model_Mean = (p_gomp+dsm_cum)/2,
                        Gompertz_Model = p_gomp,
                        DSM_Model = dsm_cum)
# Save models into a file
d = data.frame(
  date=c(Sys.Date()),
  day=c(nrow(DE)),
  country=c(country),
  first_case_date=c(min_d),
  cum_cases=c(last_confirmed),
  new_cases=c(last_delta),
  alpha_i=c( round(doubleSigmoidalModel$slope1Param_Estimate, digits=2) ),
  alpha_d=c( round(doubleSigmoidalModel$slope2Param_Estimate, digits=2) ),
  t_i=c(round(doubleSigmoidalModel$midPoint1_x, digits=0)),
  t_d=c(round(doubleSigmoidalModel$midPoint2_x, digits=0)),
  t_max=c( round(doubleSigmoidalModel$maximum_x, digits=0) ),
  I_max_d=c( round(doubleSigmoidalModel$reachMaximum_y, digits=0) ),
  I_final_d=c( round(doubleSigmoidalModel$endDeclinePoint_y, digits=0) ),
  I_max_c=c( round(I_max_c, digits=0) ),
  alpha_c=c( alpha_c ),
  beta_c=c( beta_c ),
  fc_0=c(Prediction$Model_Mean[1]),
  fc_1=c(Prediction$Model_Mean[2]),
  fc_2=c(Prediction$Model_Mean[3]),
  fc_3=c(Prediction$Model_Mean[4]),
  fc_4=c(Prediction$Model_Mean[5]),
  fc_5=c(Prediction$Model_Mean[6]),
  fc_6=c(Prediction$Model_Mean[7]),
  fc_7=c(Prediction$Model_Mean[8]),
  t_converge=c(convergent),
  assessment=c(assessment)  
)
write.table(d, file = "../models/history.csv", sep = ",", col.names = F, row.names=F, append = T, qmethod = "double")

```

---
title: "Projections for Covid-19 cases in `r countryname`"
author: Raphael Volz (rv@volzinnovation.com)
date: Last updated on `r Sys.Date()`
output: bookdown::html_document2
---

_**Executive Summary:** We expect a total of `r format(round(I_max_c, digits=0), big.mark = ",", digits=0)` cases in the first wave of the COVID-19 pandemic in `r countryname`, this is an additional `r format(ifelse((I_max_c - last_confirmed > 0),(I_max_c - last_confirmed),0), big.mark=",")` cases compared to current case total. New cases have peeked on `r DE$Date[t_max][1]` with `r format(I_max_d, big.mark=",")` new cases. Our models assume that after end of the current wave around `r format(I_final_d, big.mark=",")` new cases will appear on an average day, since SARS-CoV-2 is now endemic. Quick, effective, and differentiated mitigation measures need to be in place to prevent a new outbreak._


# New Cases

According to the Johns Hopkins University global COVID-19 data set [1] a total of `r format(last_confirmed, big.mark=",")` COVID-19 cases were reported in `r countryname` on `r yesterday`. This number of cases increased by `r round(last_increase_percent,digits=1)` percent over the cumulative total cases known one day earlier. At this rate of cases double every `r round(double_days, digits=0)` days and the basic reproduction rate $R_0$ in `r countryname` can be estimated at `r round(last_R0,digits=1)`, an estimate of the average number of other people infected by a single infected person. Figure \@ref(fig:daily-plot) shows the time line of new cases reported in the country (black points) together with a five day moving average (purple line) that smoothes daily fluctuations in the reporting that are due to fewer capacities on weekends, holidays and other reporting mishaps.

```{r daily-plot, fig.cap="New cases and five day moving average.", echo=FALSE}
# plot daily deltas
plot(delta,
#           main=paste0("Daily increase in COVID-19 cases in ", country),
#           sub=paste ( "Last Data:", today, " Source: [1]" ),
           ylab="New cases",
           xlab="Date",
           type="p")
# plot week lines
for (date in index(z)) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey")
}
# plot trend line
lines(trend, col="purple")

```

## Model for new cases (DSM)

New cases peek after a certain time of growth that follows a sigmoidal function and then recedes to a lower level as the epidemic wave fades out and enters an endemic state. This behavior can be modelled with a double-sigmoid function $f_{base}$, that is obtained by the product of two sigmoidal functions [2].

$$ f_{base}(t) = \frac{1}{ 1+e^{-\alpha_{i} * (t - t_{i})} } * \frac{1}{ 1+e^{-\alpha_{d} * (t - t_{d})} }$$

This function has a maximum $f_{max} = max(f_{base}(t_{max}))$ at $t_{max}$.

The model can obtain predictions related to an observed maximum of new cases $I_{max_{d}}$ and an endemic number of daily cases $I_{final_{d}}$ as an asymptote (that is ideally 0). Let $f_{d}(t)$ be a piecewise function that switches between the growth and decay phase at $t_{max}$, such that

$$\begin{aligned}f_{d}(t) = \left\{
\begin{array}{cc} 
                \frac{I_{max_{d}}}{f_{max}} * f_{base}(t)  & \hspace{5mm} t \leq t_{max} \\
                \frac{I_{max_{d}} - I_{final_{d} }}{f_{max}} * f_{base}(t) + I_{final_{d}} & \hspace{5mm} t > t_{max} \\
\end{array} \right.
\end{aligned}$$

where $t_{i}$ is the midpoint and $\alpha_{i}$ the slope of the increase of cases and $t_{d}$ is the midpoint and $\alpha_{d}$ is the slope of the decrease in cases. $I_{max_{d}}$ the peak of new cases and $I_{final_{d}}$ the asymptotic final value of new cases at the end of the infections wave. Figure \@ref(fig:daily-model) shows the current _double-sigmoidal model_ ( orange line) that is fitted for the new case time series of `r countryname`, where the vertical dashed line highlights the midpoint $t_i$ of the growth phase, the vertical solid line indicates the peak of new infections $t_max$ and the vertical dotted line the midpoint $t_d$ of the decay phase. The dashed horizontal line marks the estimated asymptote of endemic new cases $I_{final_{d}}$ as this infection wave ends. 

```{r daily-model, fig.cap="Double sigmoidal model (DSM) for new cases.", echo=FALSE}

# Plot model preditions
plot(DE$Date,new_dsm, type="l", col="orange", lwd=2, ylab="Double Sigmoidal model", xlab="Date", ylim=c(0, max(new_dsm,coredata(delta)) + 10 ))
# plot Week Lines
for (date in DE$Date) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey", lty=3)
}
# Add lines for key dates from DSM model
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1], lty=2, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1], lty=1, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1], lty=3, col="orange")
# add asymptote I_final_d
abline(h=doubleSigmoidalModel$endDeclinePoint_y, lty=2, lwd=2, col="orange")
# add points for reported data
points(DE$Date[seq(2,nrow(DE))],coredata(delta))

```

## Reliability of the DSM

Since the model requires six parameters, it also requires more data points. Hence, it is only feasible to fit this model in later phases of the pandemic (ideally after the peak of daily infections occured) and the model is highly sensitive to new data, particularly with respect to finding the final asymptote$I_{final_{d}}$as the following comparison of the model shows. Nevertheless the model renders interesting characteristica of the past developments.

| Model for Daily Increases         | Current Estimate                                                                                                          | Last Week Estimate                                                                                         | Delta (absolute)                                                                              |
| :-------------------------------- | -------------------------------------------------------------------------------------------------------------------------- | ------------------------------------------------------------------------------------------------------------- | --------------------------------------------------------------------------------------------- |
| $I_{max_{d}}$ Peak new cases          | `r round(doubleSigmoidalModel$reachMaximum_y, digits=0)`                                                                   | `r round(dsm_ago$reachMaximum_y, digits=0)`                                                                   | `r round(doubleSigmoidalModel$reachMaximum_y - dsm_ago$reachMaximum_y, digits=0)`             |
| $I_{final_{d}}$ Endemic new cases          | `r round(doubleSigmoidalModel$endDeclinePoint_y, digits=0)`                                                                | `r round(dsm_ago$endDeclinePoint_y, digits=0)`  | `r round(doubleSigmoidalModel$endDeclinePoint_y - dsm_ago$endDeclinePoint_y, digits=0)`       |
| $\alpha_{i}$ Slope of growth phase    | `r round(doubleSigmoidalModel$slope1Param_Estimate, digits=2)`                                                             | `r round(dsm_ago$slope1Param_Estimate, digits=2)`                                                             | `r round(doubleSigmoidalModel$slope1Param_Estimate - dsm_ago$slope1Param_Estimate, digits=2)` |
| $t_{i}$ Midpoint of growth phase   | `r DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1]` (Day `r round(doubleSigmoidalModel$midPoint1_x, digits=0)`) | `r DE$Date[round(dsm_ago$midPoint1_x, digits=0)][1]` (Day `r round(dsm_ago$midPoint1_x, digits=0)`)              | `r round(doubleSigmoidalModel$midPoint1_x - dsm_ago$midPoint1_x, digits=0)`                   |
| $t_{max}$ Peak day                   | `r DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1]` (Day `r round(doubleSigmoidalModel$maximum_x, digits=0)`)     | `r DE$Date[round(dsm_ago$maximum_x, digits=0)][1]` (Day `r round(dsm_ago$maximum_x, digits=0)`)                  | `r round(doubleSigmoidalModel$maximum_x, digits=0) - round(dsm_ago$maximum_x, digits=0)`      |
| $\alpha_{d}$ Slope of decay phase       | `r round(doubleSigmoidalModel$slope2Param_Estimate, digits=2)`                                                             | `r round(dsm_ago$slope2Param_Estimate, digits=2)`                                                             | `r round(doubleSigmoidalModel$slope2Param_Estimate - dsm_ago$slope2Param_Estimate, digits=2)` |
| $t_{d}$ Midpoint of decay phase | `r DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1]` (Day `r round(doubleSigmoidalModel$midPoint2_x, digits=0)`) | `r DE$Date[round(dsm_ago$midPoint2_x, digits=0)][1]` (Day `r round(dsm_ago$midPoint2_x, digits=0)`)              | `r round(doubleSigmoidalModel$midPoint2_x, digits=0) -round(dsm_ago$midPoint2_x, digits=0)`   |

In order to address the shortcoming of model instability under new data `r ifelse(I_max_dsm>0, " and the situation that no decay could be detected yet","")`, we can attempt to fit a different model working with cumulative case numbers instead.

# Cumulative cases

The cumulative total of COVID-19 cases in `r countryname` is the number of all COVID-19 cases known up to yesterday and **includes all patients already recovered from the infection**. This corresponds to the _Infected (I)_ number in the standard epidemiological SIR model. Figure \@ref(fig:cumulative-plot) shows the data points of the time series together with a Gompertz model for the cumulative cases (red line) as well as the cumulated double-sigmoidal model (orange line). The dashed horizontal lines show the expected maximum number of cases of the `r ifelse(I_max_dsm>0,"respective models", "Gompertz model ( the cumulated double-sigmoidal model does not have an asymptote )")`. `r ifelse(I_final_d>0 && I_max_dsm>0 ,paste0(" _Note:_ The cumulated double-sigmoidal model does not have an asymptote, since $I_{final_{d}}$ is not zero, the line can be considered as the number of infections when the infective wave ends and the endemic state starts. "),"" )`


```{r cumulative-plot, fig.cap="Cumulative cases with DSM and Gompertz models", echo=FALSE}
# Plot Infections over Time
plot(DE$Confirmed ~ DE$Date,
#     log="y", # Logarithmic Scale
#     main=paste0("COVID-19 cases in ", country),
     sub=paste ( "Last Data:", today, " JHU Github repo" ),
     ylab="Total Cases",
     xlab="Date",
     ylim=c(0,max(I_max_dsm, I_max_c, DE$Confirmed)+100)
     )
# plot Week Lines
for (date in DE$Date) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey", lty=3)
}
# Add lines for key dates from DSM model
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1], lty=2, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1], lty=1, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1], lty=3, col="orange")
# plot Gompertz model
lines(DE$Date, new_gomp, col="red", lwd=2)
# plot DSM model
lines(DE$Date, new_dsm_cum, col="orange", lwd=2)
# plot cumulative confirmed infections
points(DE$Date, DE$Confirmed, col="black")
# add asymptotes I_max
abline(h=I_max_c, lty=2, lwd=2, col="red")
# add asymptote I_max_dsm, only if there is already some recovery
if(I_max_dsm > 0) { abline(h=I_max_dsm, lty=2, lwd=2, col="orange") }

```

## Gompertz model for cumulative cases

The Gompertz model [3] is a sigmoid model that is often fitted to growth data, and often prefered over the standard logistic model (als called Verhulst model) due its asymetry. The Gompertz can be considered as a special case of the four parameter Richards model and is computationally easier to fit as it has one parameter less. The model is based on the cumulative distribution function $f_{c}(t)$, which is a three parameter logistic function:

$$ f_{c}(t) = I_{max_{c}}  * e^{ - \beta_{c} * e^{ - \alpha_{c} * t }} $$

Where the parameter $I_{max_{c}}$ is the estimated maximum number of cumulative cases (asymptote), $\beta_{c}$ inflects the curve along time, and $\alpha_{c}$ is the maximum absolute growth rate of cumulated cases.

## Reliability of the Gompertz model

The Gompertz model change over time as new data is available. Obviously, we are interested whether the models are reliable and in particular show some convergence. When we compare the current model with the model calculated using the information that was available last week, we can again observe how the model parameters have changed.

| Parameter               | Current Estimate                             | Last Week Estimate                             | Delta (absolute)                                                             |
| :---------------------- | --------------------------------------------- | ------------------------------------------------- | :--------------------------------------------------------------------------- |
| $I_{max_{c}}$ Total number of cases | `r round(I_max_c, digits=0)`                  | `r round(gomp_ago$m$getAllPars()["a"], digits=0)`                  | `r round(I_max_c, digits=0)  - round(gomp_ago$m$getAllPars()["a"], digits=0)`                |
| $\alpha_{c}$ Growth rate      | `r alpha_c`                  | `r gomp_ago$m$getAllPars()["c"]`                  | `r alpha_c - gomp_ago$m$getAllPars()["c"]`                  |
| $\beta_{c}$ Time shift       | `r round(beta_c, digits=0)` | `r round(gomp_ago$m$getAllPars()["b"], digits=0)` | `r round(beta_c - gomp_ago$m$getAllPars()["b"], digits=0)` |

# Projections

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}

```

## Methodology

Putting in future dates into the current models allows to project future cases. _Note:_ The actual development can differ from model predictions, for example if social distancing norms change or methods in data collection are altered. If such changes happen, a model fitted to the past is no longer valid.


```{r nowcast, fig.cap="Nowcast of past models for yesterday",  echo=FALSE}

# Plot Nowcast
plot(comparison$Date, comparison$gomp,
     # main="Past Predictions for Today", 
     ylab=paste0("Predicted cum. cases on ", yesterday), 
     xlab="Gompertz model based on cases known on date",
     type="l",
     ylim=c(0,DE$Confirmed[nrow(DE)]*1.5),
     col="red"
     )
# Shade the convergent area
if(convergent> 0) {
  rect(DE$Date[convergent],-1000000,max(DE$Date)+10,1000000000,col="lightblue",border=0)
}
if (convergent == 0) {
  assessment = FALSE
}
# Lines
lines(comparison$Date, comparison$gomp, col="darkblue", lwd=2)
# Plot last known confirmed cumulative value (dashed line)
abline(h=DE$Confirmed[nrow(DE)], lty=2)
# plot Week Lines
for (date in DE$Date) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey", lty=3)
}
# Add lines for key dates from DSM model
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1], lty=2, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1], lty=1, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1], lty=3, col="orange")
# Add line when model converged
# (v=DE$Date[convergent], lty=2, lwd=2, col="green")
```

## Nowcast

Yesterday `r DE$Confirmed[nrow(DE)]` COVID-19 cases were known in `r countryname` [1]. Figure \@ref(fig:nowcast) illustrates how past Gompertz models (red line) have predicted the last known number of cumulative COVID-19 cases (dashed black line). `r ifelse(convergent>0, paste0("The lightblue shaded area indicates that the Gompertz model of ",DE$Date[convergent]," (day ",convergent,") and all later models could predict the case number for ", yesterday, " with at most  ",criteria," percent difference to the actual case numbers. Hence, the Gompertz models could reliably predict the actual case numbers for the coming ", (nrow(DE)-convergent), " days."), "Unfortunately the Gompertz models did not yet converge and cannot provide reliable nowcasts and **we must regard all forecasts with great scepticism.**")`

```{r, echo=FALSE}
# Plot prediction
#plot( p_gomp ~ Date,
#     # log="y", # Logarithmic Scale
#     type="l",
#     main=paste("7 day Forecast of COVID-19 cases in", country),
#     sub=paste ( "Forecast on based on JHU data" ),
#     ylab="COVID-19 cases",
#     lwd=2,
#    col="red",
#xlab="Date",
#)
# plot known data points
#points(DE$Date, DE$Confirmed, col="black")
```

## Forecast

Table \@ref(tab:forecast) presents the predicted number of total COVID-19 cases for the upcoming week that are predicted using our models.

```{r forecast, echo=FALSE}
# Print result table
kable(Prediction,
      digits = c(0,0,0,0),
      format.args = list(big.mark = ",",scientific = FALSE),
      caption=paste("Projected COVID-19 Cases in", country),
     col.names = c("Date", "Model Mean", "Gompertz Model", "DSM Model"))
```

# Discussion and outlook

Epidemiological models are useful for policy makers, e.g. to assess needed ICU beds, required (additional) restrictions of daily life, etc. However models require (often disputed) assumptions for model parameters. The data-driven approach demonstrated here avoids parameter assumptions and uses past data to fit models that not only describe the past developments but provide reliable forecasts for the nearer future of an epidemic. Next to policy makers, the general public can use the projections to assess their expectations when business or schools might reopen (or reclose).

As a next step data about the population and population structure could be used to compare between countries, for example to assess the effectiveness of different public policies for controlling the disease.

# References

1.  Johns Hopkins University, _COVID-19 global case time series (csse_covid_19_time_series)_, online https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/ (Last accessed `r Sys.Date()`)
2.  Mehmet Caglar, Ashley Teufel, Claus Wilke, _Sicegar: R package for sigmoidal and double-sigmoidal curve fitting_ in PeerJ. 2018 Jan 16;6:e4251. doi: 10.7717/peerj.4251. eCollection 2018.
3.  Benjamin Gompertz, _On the nature of the function expressive of the law of human mortality and a new mode of determining the value of life contingencies_. Philosophical Transactions of the Royal Society of London : Biological Sciences. 1825, 182:5 13-85.

# (APPENDIX) Appendix {-}

# Gompertz model history

The following table states the known COVID-19 infections and duration of the pandemic in `r countryname` on a given day. It also states all _three parameters_ of the model estimated with data available on that date and reports the root mean square error (RMSE) of the model on the available data RMSE (Training). The information made available after model creation is used as a test of the model precision on unknown data and reported as RMSE (Test). The forecast of the given model for yesterday is provided in the last column.

```{r, echo=FALSE}
# Print result table
kable(models2,
      digits = c(0,0,0,0,2,2,0,0,0),
      format.args = list(big.mark = ",",scientific = FALSE),
      col.names = c("Date", "# Cases", "# Days", "I_max", "beta", "alpha", "RMSE (Train)", "RMSE (Test)", "Nowcast"))

```

# Parameter convergence of the Gompertz model

Figures \@ref(fig:evol-Imax) - \@ref(fig:evol-beta) present how the model parameters change over time as new cases are reported every day.


```{r evol-Imax, fig.cap="Parameter evolution of expected maximum (I_max_c)", echo=FALSE}
# Plot parameter evolution for I_maxs
plot(models2$Date, models2$Maximum,
     ylab="Expected maximum",
     xlab="Gompertz model based on cases known on date",
     type="l",
     ylim=c(0,max(DE$Confirmed)*4)
    )
# plot Week Lines
for (date in DE$Date) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey", lty=3)
}
# Add lines for key dates from DSM model
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1], lty=2, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1], lty=1, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1], lty=3, col="orange")
```

```{r evol-alpha, fig.cap="Parameter evolution of growth rate (alpha_c)", echo=FALSE}
# Plot parameter evolution for alphae
plot(models2$Date, models2$C,
     ylab="Alpha",
     xlab="Gompertz model based on cases known on date",
     type="l",
     ylim=c(0,models2$C[nrow(models2)]*4)
    )
# plot Week Lines
for (date in DE$Date) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey", lty=3)
}
# Add lines for key dates from DSM model
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1], lty=2, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1], lty=1, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1], lty=3, col="orange")
```

```{r evol-beta, fig.cap="Parameter evolution of time shift (beta_c)", echo=FALSE}
# Plot parameter evolution for betae
plot(models2$Date, models2$B,
     ylab="Beta",
     xlab="Gompertz model based on cases known on date",
     type="l",
     ylim=c(0,models2$B[nrow(models2)]*4)

    )
# plot Week Lines
for (date in DE$Date) {
  if( as.numeric(strftime(as.Date(date, origin="1970-01-01"), "%u")) == 7 )
    abline(v=as.Date(date, origin="1970-01-01"), col="grey", lty=3)
}
# Add lines for key dates from DSM model
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint1_x, digits=0)][1], lty=2, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$maximum_x, digits=0)][1], lty=1, col="orange")
abline(v=DE$Date[round(doubleSigmoidalModel$midPoint2_x, digits=0)][1], lty=3, col="orange")
```

# Source code for model estimation

The source code for estimating the models and computing metrics, as well as generating this report can be found on [Github](https://github.com/volzinnovation/covid-19_SARS-CoV-2_corona/blob/master/src/COVID-19.rmd) and be run with the [free software environment for statistical computing and graphics R](https://www.r-project.org/) that is available for Windows, Mac and Linux.